---
title: "Quick Draft Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
editor_options: 
  chunk_output_type: console
runtime: shiny
---

```{r setup, include=FALSE}

library(flexdashboard)
library(tidyverse)
library(plotly)
library(ggtext)
library(shiny)

```



```{r global, include = FALSE}

# load data
df <- read_csv("ff2023_merged_data.csv") 

vars <- c("Pick Number",   
          # "DraftSheets Position Rank", 
          "FantasyPros Rank", "FantasyPros Projected Points",
          "DraftSheets Rank", "DraftSheets VAL", 
          "ADP (2QB)")

pos.list <- c("QB", "WR", "RB", "TE", "D/ST", "K")
pos.list2 <- c("QB", "WR", "RB", "Flex", "TE", "D/ST", "K")

starter.num <- data.frame(Position = pos.list2,
                          num = c(2, 3, 2, 1, 1, 1, 1))

colors7 <-  c("#CC79A7", "#56B4E9", "#009E73", "#F0E442", 
              "#0072B2", "#D55E00", "#E69F00")
pos.colors <- colors7
names(pos.colors) <- pos.list2

var.switch <- function(var){
  switch(var,
         "Pick Number" = "Pick",
         "ADP (2QB)" = "ADP",
         "DraftSheets VAL" = "val",
         "DraftSheets Rank" = "val_rank",
         # "DraftSheets Position Rank" = "pos_rank", 
         "FantasyPros Rank" = "ECR", 
         "FantasyPros Projected Points" = "FPTS"
         )
}


# processing for page 2
# identifying starters and bench
df2 <- df %>%
  left_join(starter.num) %>%
  # mutate(Position = factor(Position, levels = pos.list2 )) %>%
  group_by(Manager, Position) %>%
  arrange(Manager, Position, desc(FPTS)) %>%
  mutate(starter = ifelse(FPTS >= FPTS[num], "Starter", "Bench"))

df.flex <- df2 %>%
  filter(starter == "Bench", Position %in% c("WR", "RB", "TE")) %>%
  group_by(Manager) %>%
  slice_max(FPTS, n = 1) %>%
  mutate(Position2 = "Flex") %>%
  select(Player, Position2)

df3 <- df2 %>%
  left_join(df.flex) %>%
  mutate(Position2 = ifelse(is.na(Position2), Position, Position2),
         Position = factor(Position, levels = pos.list2 ),
         Position2 = factor(Position2, levels = pos.list2 ),
         starter = ifelse(Position2 == "Flex", "Starter", starter)) %>%
  ungroup() %>%
  arrange(Manager, desc(starter), Position, desc(FPTS))

```



```{r reactive}

# define reactive elements for shiny
  # page 1:
  x <- reactive(input$x)
  y <- reactive(input$y)
  positions <- reactive(input$positions)
  facet <- reactive(input$facet)
  
  # page 2:
  criteria <- reactive(input$criteria)
  startSelect <- reactive(input$startSelect)
  positions2 <- reactive(input$positions2)
  

```

# Draft Analysis

## Sidebar {.sidebar data-width=160}
<!-- #### **Person 1** -->
```{r sidebar1}

# sidebar is the UI elemennt that contains the inputs for reactive stuff

selectInput(inputId = "x",
            label = "x-axis",
            choices = vars,
            selected = vars[1],
            selectize = TRUE,
            )

selectInput(inputId = "y",
            label = "y-axis",
            choices = vars,
            selected = vars[2],
            selectize = TRUE,
            )

selectInput(inputId = "positions",
            label = "Position Filter",
            choices = pos.list,
            selected = pos.list,
            multiple = TRUE,
            selectize = FALSE,
            size = 6,
            )

radioButtons(inputId = "facet",
             label = "Separate panels for each team?",
             choices = c("No", "Yes"),
             selected = "No")


```


##
```{r }

x2 <- reactive(var.switch(x()))
y2 <- reactive(var.switch(y()))

fillCol(height = 1000, flex = 1, 
  plotlyOutput("g1", height = "100%")
)

output$g1 <- renderPlotly({

  g <- df %>%
    filter(Position %in% positions()) %>%
  ggplot(aes(.data[[x2()]], .data[[y2()]], color = Manager)) +
    geom_point(aes(text = Player)) +
    theme_bw(base_size = 16) +
    theme(legend.title = element_blank()) +
    labs(x = x(),
         y = y())

  if(all(c(x(), y()) %in% c("Pick Number",  
          "FantasyPros Rank",
          "DraftSheets Rank",
          "ADP (2QB)"))){
    g <- g +
      geom_abline(aes(slope = 1, intercept = 0))
  } else{
    g <- g +
      geom_smooth(color = "gray50", se = FALSE)
      
  }
  
  if(facet() == "Yes"){
    g <- g + 
      facet_wrap(vars(Manager), nrow = 2) + 
      theme(legend.position = "none")
  }

  ggplotly(g)

})



```

# Team Projections and Value
## Sidebar {.sidebar data-width=160}
<!-- #### **Person 1** -->
```{r sidebar2}

# sidebar is the UI elemennt that contains the inputs for reactive stuff

radioButtons(inputId = "criteria",
             label = "Projections/VAL",
             choices = c("FantasyPros Projected Points", "DraftSheets VAL"),
             selected = "FantasyPros Projected Points")

radioButtons(inputId = "startSelect",
             label = "Starters/Bench",
             choices = c("Starters", "Bench", "Both"),
             selected = "Starters")

selectInput(inputId = "positions2",
            label = "Position Filter",
            choices = pos.list2,
            selected = pos.list2,
            multiple = TRUE,
            selectize = FALSE,
            size = 7,
            )

```


##
```{r}
# # for testing outside of running shiny  
# dfpoints <- df3 %>%
#   filter(starter == "Starter") %>%
#   group_by(Manager) %>%
#   summarise(total_points = sum(FPTS)) %>%
#   arrange(desc(total_points))
# 
# df3 %>%
#   left_join(dfpoints) %>%
#   # drop_na(val) %>%
#   filter(starter != "Bench") %>%
#   mutate(Manager = fct_reorder(Manager, total_points),
#          Position = fct_rev(Position2),
#          Player = fct_reorder(Player, FPTS)) %>%
#   ggplot(aes(Manager, FPTS, text = Player)) +
#   geom_bar(aes(fill = Position, alpha = starter), stat = "identity") +
#   geom_text(data = dfpoints %>% mutate(Player = ""),
#             aes(y = total_points, label = round(total_points)), hjust = -.1) +
#   scale_y_continuous(expand = expansion(mult = c(.05, .1))) + 
#   scale_alpha_manual(values = c(.6, 1)) +
#   coord_flip() +
#   theme_bw()
# 
# ggplotly(tooltip = c("x", "y", "fill", "text"))

```

```{r}


fillCol(height = 1000, flex = 1, 
  plotlyOutput("g2", height = "100%")
)

output$g2 <- renderPlotly({
  
  if(startSelect() == "Starters"){
    df3 <- df3 %>%
      filter(starter == "Starter")
  } else if(startSelect() == "Bench"){
    df3 <- df3 %>%
      filter(starter == "Bench")
  }
  
  df3 <- df3 %>%
    mutate(Position = fct_rev(Position2)) %>%
    filter(Position %in% positions2())
  
  if(criteria() == "FantasyPros Projected Points"){
    df_sort <- df3 %>%
      group_by(Manager) %>%
      summarise(total = sum(FPTS)) %>%
      arrange(desc(total)) %>%
      mutate(location = total + max(total)*.05,
             label = round(total))
  } else if(criteria() == "DraftSheets VAL"){
    df_sort <- df3 %>%
      group_by(Manager) %>%
      summarise(total = sum(val, na.rm = TRUE),
                location = sum(val[val>0], na.rm=TRUE)) %>%
      arrange(desc(total)) %>%
      mutate(location = location + max(.7, max(location) * .05),
             label = paste0("Net =\n", round(total, 1)))
  }
  
  criteria2 <- var.switch(criteria())
  
  g <- df3 %>%
    left_join(df_sort) %>%
    drop_na(.data[[criteria2]]) %>%
    mutate(Manager = fct_reorder(Manager, total),
           Player = fct_reorder(Player, .data[[criteria2]])) %>%
    ggplot(aes(Manager, .data[[criteria2]], fill = Position, text = Player)) + 
      geom_bar(aes(alpha = starter), stat = "identity") + 
      geom_text(data = df_sort %>% mutate(Player = ""),
                aes(y = location, label = label, fill = NULL)) +
      geom_hline(yintercept = 0) + 
      scale_y_continuous(expand = expansion(mult = c(.05, .1))) + 
      scale_alpha_manual(values = 1) + 
      scale_fill_manual(values = pos.colors) + 
      coord_flip() + 
      theme_bw() + 
      labs(fill = NULL,
           alpha = NULL)
  
  if(startSelect() == "Both"){
    g <- g +
      scale_alpha_manual(values = c(.6, 1))

  }

  ggplotly(g, tooltip = c("x", "y", "fill", "text"))

})

```



# Data Sources

FantasyPros superflex rankings (which is basically 2QB), downloaded 2023-09-06:[https://www.fantasypros.com/nfl/rankings/ppr-superflex-cheatsheets.php](https://www.fantasypros.com/nfl/rankings/ppr-superflex-cheatsheets.php)

FantasyPros Projections, downloaded 2023-09-06: [https://www.fantasypros.com/nfl/projections/flex.php?week=draft&scoring=PPR&week=draft](https://www.fantasypros.com/nfl/projections/flex.php?week=draft&scoring=PPR&week=draft)

DraftSheets, 2023-09-03 update: [https://old.reddit.com/r/fantasyfootball/comments/166bxuk/another_draft_sheet_option_update_8312023/](https://old.reddit.com/r/fantasyfootball/comments/166bxuk/another_draft_sheet_option_update_8312023/)

More info on DraftSheets: [https://old.reddit.com/r/fantasyfootball/comments/15jexnu/another_draft_sheet_option/?ref=share&ref_source=link](https://old.reddit.com/r/fantasyfootball/comments/15jexnu/another_draft_sheet_option/?ref=share&ref_source=link)

- basically VAL is just data pulled from FantasyPros and adjusted for a "baseline player" at each position

ADPs, downloaded 2023-09-06: [https://fantasydata.com/nfl/2qb-adp](https://fantasydata.com/nfl/2qb-adp)

